---
title: "Fonts in Windows"
author: "Pep Porrà"
date: "2022-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(extrafont)
library(ggplot2)
```

## Goal

Explore how to use all fonts available in Windows OS in R

## Context

Once you have the new TTF font installed in windows (using the OS function for it [microsoft](https://support.microsoft.com/en-us/windows/how-to-install-or-remove-a-font-in-windows-f12d0657-2fc8-7613-c76f-88d043b334b8), use the `extrafont` package to install the font in R.

## Installing new font in windows

* Download font (TTF file). Example: [Google fonts](https://fonts.google.com/specimen/Josefin+Sans?query=josefin&preview.text=jose&preview.text_type=custom)
* Double click in the TTF font within the ZIP file downloaded
* Click Install button

## {extrafont} package

Process to use a new font in R in windows using package {extrafonts}:

1. Install first the font using windows operating system (reference given above)
1. Check where the folder where the font was installed
1. Import the font to R (using pattern parameter) with `import_font()`. 
You will have to answer yes to the question on the console if prompt is not
set to FALSE
1. Load font into win device (or any other you want to use, like pdf)

Reference: [Medium](https://medium.com/analytics-vidhya/5-easy-steps-to-import-new-fonts-into-r-6651bf263a07)

### Example

I installed Josefin Sans from [Google fonts](https://fonts.google.com/specimen/Josefin+Sans?query=josefin&preview.text=jose&preview.text_type=custom)
using windows OS procedure [microsoft](https://support.microsoft.com/en-us/windows/how-to-install-or-remove-a-font-in-windows-f12d0657-2fc8-7613-c76f-88d043b334b8).

Second step is to check where the font was installed. We use the internal function
in extrafont package `extrafont:::ttf_find_default_path()`


```{r}
extrafont:::ttf_find_default_path()
```

With `list.files()`, we search in which folder of them the font you want 
to load to R is located

```{r}
list.files(path = "C:\\windows/Fonts", pattern = "\\.ttf") |> 
  grep(pattern = "Josefin-Sans")
```
```{r}
list.files(path = "C:\\Users\\josep\\AppData\\Local/Microsoft/Windows/Fonts", 
  pattern = "\\.ttf") |> 
grep(pattern = "Josefin", value = TRUE)
```
Alternatively, you can use {systemfonts} package:

```{r}
all_my_fonts <- systemfonts::system_fonts() 
grep(all_my_fonts$path, pattern = "Josefin", value = TRUE)  
```

Next step is to import the font to R using `extrafont::import_font()`

```{r}
extrafont::font_import(path = "C:\\Users\\josep\\AppData\\Local/Microsoft/Windows/Fonts",
  pattern = "\\.ttf", prompt = FALSE)
```

If `prompt = FALSE` is not included, you will have to answer yes to a question 
in your console.

You can check that the font is available as follows

```{r}
extrafont::fonts()
```

Last step before using the fonts is to load the fonts into the device you are
using `extrafont::loadfonts(device=”win”)`

```{r}
extrafont::loadfonts(device = "win")
```
Use `extrafont::fonts()` function to know the precise name of the font you want
to use

```{r}
extrafont::fonts()
```

```{r}
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
  ggtitle("Title text goes here") +
  theme(plot.title = element_text(size = 16, family="Josefin Sans Thin"))
```


```{r}
df <- data.frame(x = runif(10), y = rnorm(10))
plot(df$x, df$y, family="Josefin Sans Thin")
```

## Problems

In case of problems, you can check the following references

* [issue with font_import()](https://stackoverflow.com/questions/59949572/i-cannot-import-fonts-using-font-import)
* [Issue "No Font Name"](https://stackoverflow.com/questions/61204259/how-can-i-resolve-the-no-font-name-issue-when-importing-fonts-into-r-using-ext)
* [GITHub extrafonts](https://github.com/wch/extrafont)

Also, you can start from scratch by deleting all fonts in the extrafont database 
with just reinstalling the extrafontdb package

```{r, eval=FALSE}
# Run it if you want to reset extrafont database
install.packages("extrafontdb")
```

You can check that no fonts is in the database using `fonts()`

```{r}
extrafont::fonts()
```


Package version of Rttf2pt1.
It seems there is an issue with last version of Rttf2pt1 (1.3.10) 
[Issue](https://github.com/wch/Rttf2pt1/issues/10) that is solved by installing
an older version 1.3.8

Check which version you have  with

```{r}
packageVersion("Rttf2PT1")
```
and reinstall the older version with 

```{r}
library(remotes)
remotes::install_version("Rttf2pt1", version = "1.3.8")
```


